### 2011-08-12 doc/paper_ambient/paper_ambient.tex
some corrections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN PACKAGE
\documentclass[11pt]{amsart}       % AMS template just for the draft
\usepackage{geometry}                 % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                        % .letter or a4paper or a5paper or ... 
%\geometry{landscape}                 % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}     % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}                   % FIgure
\usepackage{amssymb}                 % Math symbol
\usepackage{epstopdf}                   % for working in .pdf and  not .dvi (convert .eps, .jpg,  etc to .pdf)
\usepackage{psfrag}
\usepackage[T1]{fontenc}
\usepackage{aeguill}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END PACKAGE
\graphicspath{{PLOT/SinglePDGEMM/}{PLOT/VectorPDGEMM/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN AUTHORS AND INSTITUTIONS
\title{AMBIENT; an  execution and an MPI-CUDA scheduler  }
\author[A. Kosenkov, T. Ewart, B. Bauerb, A. Kantian,  M. Troyer, T. Giarmarchy]{Alexandr Kosenkov, Timoth\'ee Ewart, Bela Bauer, Adrian Kantian, Matthias Troyer and Thierry Giamarchy }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN AUTHORS AND INSTITUTIONS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN MACRO
\makeatletter
\newcommand\etc{\textit{etc}\@ifnextchar.{}{.\@}}
\newcommand\eg{\textit{e.g. }}
\newcommand\ie{\textit{i.e. }}
\newcommand\cf{\textit{c.f. }}
\renewcommand{\labelitemi}{$\diamond$}

\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END MACRO
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN SYMBOL
\def\cloud{\includegraphics[width=0.2cm]{FIGURES/cloud}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END SYMBOL
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN INTRODUCTION
\section{Introduction}

%Nowadays, the numerical simulations get an essential position, as experiments,  for studying physical 
%properties of many problems. The numerical simulations are presented inside every fields of sciences 
%and engineering, as classical and quantum mechanic, quantum chemistry,  \etc. As usual, the users always
%want increase the size of their problems, as billion number of particles for a kinetic simulation, billion 
%of meshes for a fluid dynamic computation, or else, matrixes of million entries for a quantum mechanic,
%or the linear algebra problems. Thus, the high performance computing is the solution for solving very large problem,
%with the peta flops  and the future generation of exa flops computers, the engineers and the scientists will
%have a quasi unlimited set of resources for modeling, nevertheless, they will be confronted to a challenge, have compatible
%programs for these machines.

%A large set of programs; solvers, and software solutions are now availabled to solve large problem in a lot of fields, \eg we can 
%cite as popular in a few scientist field :  {\sc Quantum ESPRESSO} \cite{QE-2009} in quantum chemistry, Abinit \cite{ABINIT-2005, ABINIT-2009} in quantum mechanic,  ScaLAPACK \cite{SCALAPACK-1997} in linear 
%algebra,  or else SMILE in hypersonic rarefied flow \cite{SMILE-2005}. Of course, this short list is not exhaustive,
%and it can be fill up easily in every fields of science and engineering.

%If the scientists or the  engineers have the right tools, it will be necessarily  limited by the scalability of the code, in the case of the famous mathematical library ScaLAPACK, the
%efficiency  and thus the scalability, with the appropriated setting will reach up to 512 processors ($\sim  \mathcal{O}(100)$) \cite{SCALAPACK-1992}, beyond the interest is useless
%due to the degradation of the performance. In the worst case, the programmers will start  from a white page, and must design and efficient code.
%Due to the incredible number of cores, and optional accelerators as GPU, FGPA or "CELL like\footnote{Although the CELL processor is now depreciated, it herebies  
%a way of integrated processor (CPU+GPU), as the prove the last processor products of Intel (core i series) and AMD (fusion).}", the programmers 
%will have to introduce a mix programming model usually named mix-mode, based on a main MPI layer coupled with one or several additional layers \cite{EXASCALE-2010}.

%The most popular layer for mix-mode is OpenMP, and it is based on the preprocessor directives \texttt{\#pragma}, it demonstrated its performance. It is included successfully inside
%the  linear algebra libraries like the MKL from Intel or the PESSLSMP from IBM.  This ideology of the utilization of the pre-directives becomes more and more 
%important, thus it is used for vectorization of loops using the intel compiler and even for GPU with the PGI compiler. Although this ideology gives results, there is 
%a limitation for the main following reason: the creation/destruction of the threads will be extremely time consuming,
%when the  grain problem size becomes too fine \cite{SMITH-2001}. To achieve the best performance, the number of thread should be  equal to the number of cores on a socket, 
%thus the scalability will be increased at maximum by an order of magnitude.  The maximum number of processors now reach, becomes ($\sim \mathcal{O}(1000)$). 
%This ideology of implementation of the mix mode is more an evolution of the programming model than a revolution. It is an fast hand-on to implement that prolongates the life and the performance on a existing code. But,
%it will not prepare at all the code for the future, that will be automatically depreciated  by the new generation of cluster, in term of scalability.

%For getting several order of magnitude higher, we need a new programming model. The introduction of the GPU computing by NVidia a few years ago was
%a real innovation in HPC \cite{NICKOLLS-2010}. On a GPU, it is usual to program  thousand of threads. Thus, in this paper, we propose a innovative programming model to program on very large distributed cluster. We develop
%a framework named Ambient (C++) where we expend the CUDA programming ideology to  the a full cluster, thus we keep away the MPI programming model from the users.  In a way, we  extended
%the philosophy of OpenCL, OpenCL extents  the GPU programming model to a processor, assimilate a socket as a GPU, where every cores of the socket are assimilated to a streaming processor.
%We also developed inside a dynamic stack manager, that stocks and manages the execution of numerical "operations-solver" in functions of the resources of the distributed cluster.

%In this paper, we are going to describe our object oriented framework, and promote a first application based on linear algebra. Our main objective, is to 
%promote a new technology to develop  parallel application, without any interaction with the MPI and thread layer, and allow, a real access in term of performance to the next generation
%of clusters. The first possible  targets are applications that require large buffer of contiguous memory: quantum mechanic, linear algebra,  computational fluid dynamics, molecular dynamic \ldots 
%\, Until the end of this paper, we will focus on the linear algebra problems and their applications.

- Stress Performance of Ambient 
  - in terms of flexible and dynamical resource allocation
  - Stress performance boost derived from this (we need the benchmarks for that!)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END INTRODUCTION

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN WORKFLOW
\section{Ambient workflow}

Before presenting various elements of Ambient in more details, here we describe the
workflow of execution for an Ambient-parallelized application in general terms. 
As the Ambient computational architecture derives from CUDA, we will show a simplified workflow of CUDA
 vs Ambient enabled applications execution side-by-side:

\begin{table}[h]  
	\begin{center}
	 	\begin{tabular}{l l  l l} % L L  L L 
                        \hline 
                             \multicolumn{2}{c}{CUDA} &  \multicolumn{2}{c}{Ambient} \\
                             \hline
                             \multicolumn{4}{c}{  $\diamond$ Serial code}  \\
                            $\diamond$ & Kernel encountered                                  &  $\diamond$ & Kernel-pair (logistic/computa- \\
                               & & & tional) encountered \\
                               $\diamond$ &Thread block grid construction&     $\diamond$  & Kernel-pair stacked for later  \\
                                &(gridDim, blockDim)  &  &  execution  \\ 
                               $\diamond$ &  Thread blocks mapping onto de-  &  &  \\
                                & vice (GPU SMPs) &  & \\
                               $\diamond$ &  Execution of kernel: GPU SMP & & \\
                               &  executes kernel on every thread& &    \\
                               &  of corresponding thread blocks  & & \\
                               \multicolumn{4}{c}{  $\diamond$ Serial code}  \\
                               & &    $\diamond$&Stack flush signal encountered : \\
                               & & & execution of kernels \\
                                  \hline
        		 \end{tabular}  
              \caption{Workflow analogy between the execution of a CUDA and a Ambient application} \label{WORKFLOW}
	\end{center}
\end{table}

As soon as Ambient encounters the flush signal - kernel
execution is initiated along the following sequence:

\begin{itemize}
\item Execution of logistic kernel: allocation of device (MPI group) according to specified request. 
\item Execution of logistic kernel: argument's work groups decomposure according to workDim.
\item Execution of logistic kernel: selected argument's work groups mapping onto device (MPI processes).
\item Transfer of work groups to the target processes.
\item Execution of computational kernel: MPI process executes kernel thread for work items of assigned work groups.
\end{itemize}

The main reason for the difference in workflow compared to CUDA is the scale of available resources.  As practically unlimited resources
are available for a kernel's launch, the data becomes the key factor around which Ambient's design evolves.
In particular, in order to control the data movements and resource usage, we introduce the logistic kernel and make kernel calls non-blocking.

Generally, Ambient enabled application operates with abstract types that already have tiled memory structure. Upon a kernel operation 
call, the Ambient framework stacks the operation for later execution and proceeds to the following part of the serial code. This design
enables high degree of original serial code's reusage while providing powerfull instrumentation capabilities to manage resources and tune performance 
of the overall application.

When the framework encounters the signal to flush the stack of kernels - the actual execution starts. Framework's engine goes through the accumulated stack 
and executes all independent operations at once. Executing an operation first the engine invokes the associated logistic kernel and allocates specified
amount of resources for this operation. Next, the engine marks argument's tiling using specified parameters and maps resulting work groups onto allocated
processes in a user specified way. Now the engine can freely transfer the tiled data from previous owners to the allocated processes changing
the tile size if needed. After the transfer is complete the computational kernel kicks in behaving almost the same way as in CUDA with work groups
simply being mapped onto threads of MPI process as if they were GPU SMP (mainly, the difference is the user's control over processes/work groups mapping).

In the following sections, we elaborate on the role and functionalities of the different elements listed here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END WORKFLOW

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN DESCRIPTION OF THE FRAMEWORK
\section{Design of the framework}
%In this section, we describe our framework in simple terms, without  a deep technical description.
%We will focus on the main functionalities  and show the essential features.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN CUDA IDEOLOGY SECTION
\subsection{Extending the CUDA approach} \label{CUDAIDEOLOGYSECTION}

Ambient employs an extension on the CUDA approach to a distributed cluster. This is based on the observation that modern clusters employ many thousands of nodes, just like a GPU consists of thousands of small, simple cores, on each of which it runs an independent thread. In such a comparison however, the technical specifics of the basic cluster unit in terms of processor power, memory bandwidth, network speed, \etc  \, must be considered.

CPU core has much more additional capabilities comparing to the one of a GPU and therefore we can't simply downcast the CPU core to the GPU one.
Thus, the scope of the data for a single CPU (on which a single Ambient's thread typically runs) in Ambient is rather comparable to the capacity of CUDA's thread block or an OpenCL work group, the typical block size is $ 128 \times 128$. Though, the user has full control of the size and can adapt it for the needs of target problem. Like in CUDA, blocks (i.e. groups of threads) and the grid of blocks can be formed in Ambient. A basic comparison between CUDA and Ambient is shown in the table \ref{ANALOGY}.

\begin{figure}[t]
	\begin{center}
	                  %psfrag for changing the layout
	                  \psfrag{New data layout}{\tiny{new layout}}
			\includegraphics[scale=0.44]{FIGURES/MemMatrix.eps} 
	\end{center}
          \caption{Representation of the data layout inside the memory for linear algebra. We move from a single contiguous column data layout (left ) to several tile data layouts (right).}
          \label{MEMMATRIX}
\end{figure}

The work group approach imposes the adaptation of the data layout to the framework. In linear algebra,  BLAS \cite{LAWSON-1979} is \textit{de facto}  the standard for high performance dense Linear Algebra computations. It imposes column order on matrix storage, as well as a contiguous storage in memory (figure \ref{MEMMATRIX}).

For the sake of backward compatibility with the standard BLAS-based mathematical libraries (BLAS, LaPACK ScaLAPACK), we keep, the column order, but do not conserve a full contiguous buffer. As in the plasma library \cite{LTAIEF-2011}, we select
a tile data layout. It confers the ability to fit perfectly with the work group representation of Ambient, and further allows an excellent MPI/Thread control (section \ref{MPIABSTRACTION}). One of the main features of Ambient is to map this new data layout  to our work group approach, as shown in Fig. \ref{CUDAANALOGY}.

In CUDA and OpenCL, the execution model imposes kernel initialization with the appropriate resources (\eg numthread, numblock and grid dimension),
after which the user does not have to manage the memory access inside GPU memory; further, all threads can readily access any part of the GPU. Due to memory distribution in clusters we need to adapt this method of execution. Thus, we split this execution into two parts.

First, we prepare the memory, and then execute our "numerical" operation \eg a matrix multiplication in the following sequence \footnote{From here onwards, we will illustrate the execution model through parallel matrix addition and multiplication.}:
\begin{enumerate}
	\item A logistic kernel; where we select the number of processors, and map the memory with the datas.
	\item A computational kernel; where we execute the associated functions (the kernel in CUDA).
\end{enumerate}

Thus, every operations will have an associated logistic and computational kernel \eg multiplication, addition, or more sophisticated linear algebra operations

\newpage

\begin{figure}[!t]
	\begin{center}
           	        \psfrag{Ambient representation}{\tiny{\hspace{-0.4cm}Ambient representation}}
           	        \psfrag{CUDA/OpenCL representation}{\tiny{\hspace{-0.9cm}CUDA/OpenCL representation}}
	                 \psfrag{Associated mapping of the data with Ambient}{\tiny{\hspace{+0.15cm}Associated data mapping}}
	                 \psfrag{1x1 workgroup}{\tiny{\hspace{-0.5cm} \vspace{-0.1cm}1x1 workgroup}}
	                  \psfrag{of Ambient}{\tiny{\hspace{-0.4cm}\vspace{-0.9cm} of Ambient}}
	                  \psfrag{Workgroup of 4X4 threads}{\tiny{\hspace{-0.5cm} \vspace{-0.1cm}4x4 workgroup}}
	                  \psfrag{of CUDA}{\tiny{\hspace{-0.25cm}of CUDA}}
	                  \psfrag{Scope of GPU thread (left)}{\tiny{\hspace{-0.25cm}GPU scope left}}
	                  \psfrag{CPU (right)}{\tiny{\hspace{-0.7cm}CPU scope right}}       
		        \includegraphics[scale=0.44]{FIGURES/CUDAanalogy.eps} 
	\end{center}
          \caption{Corresponding of the CUDA representation, to the ambient representation, in terms of the scope of the CPU with the associated data.}
          \label{CUDAANALOGY}
\end{figure}
\begin{table}[h]
	\begin{center}
	 	\begin{tabular}{c  c  c}
		 	                                                      & CUDA-OpenCL                                & Ambient            \\  \hline 
			 Device                                        & Graphic card/Processor (socket)   & Cluster               \\  
			  Entity  of Calculation               & GPU-Thread                                      & CPU-Thread     \\  
			 \begin{tabular}{c}
			 Scope of the \\ 
			 entity calculation  
                             \end{tabular}
                             &  Buffer entry                            & Full buffer \\
			 Execution model                       & Kernel                                                &  \begin{tabular}{c}  Logistic and \\ computational kernel   \end{tabular} \\
			 work group-ID                           &  \texttt{blockIdx}                                 & \texttt{get\_block\_id(A)}   \\  
			 work group size                        &  \texttt{blockDim}                               & \texttt{get\_mem\_dim(A)}      \\  
			 grid size                                     &  \texttt{gridDim}                                  & \texttt{get\_grid\_dim(A)}  \\ \hline

		\end{tabular}
		\caption{Analogy between our framework and  the CUDA and OpenCL ideology.  \texttt{A} is the associated matrix to the  logistic and computational kernel.  
		The access of the cartesian values are done by the classical operator C/C++ \guillemotleft.\guillemotright     \, \eg \texttt{get\_mem\_dim(A).x}  	\label{ANALOGY}}
	\end{center}
\end{table}

\newpage

\subsubsection{The logistic kernel} The role of the logistic kernel is to initialize both memory and data for execution. It consists of creating a MPI-group and mapping the data to the group with the appropriate pattern (1D or 2D block-cyclic distribution for a matrix, or alternatively
a row/column-wise block distribution for two-dimensional Finite Difference Methods in fluid dynamics).  The creation of the group is done by an SQL-like request, and the data is mapped into this newly created group using a modular assignment
function. Both these functionalities are described in section \ref{SQLSECTION}.

In linear algebra, the optimal distribution in terms of load balancing is the 2D block-cyclic distributions, esp. in ScaLAPACK, but this will generally depend of the solver used. For example, in a QR factorization \cite{DEMMEL-2008} it can be demonstrated that a row-wise distribution is superior to a 2D block-cyclic decomposition for matrices with more rows than columns. Switching from one distribution scheme to another, which is a challenge in ScaLAPACK, is easily accomplished with Ambient.
 \begin{table}[h]
	\begin{center}
	 	\begin{tabular}{l l }
                           \hline 
                           \textbf{Logistic kernel }  Algorithm of the addition logistic kernel  &    \\ \hline
                           \textbf{Require}                 A class Matrix  that fits with Ambient framework  &  \\
                           \begin{tabular}{c l} 
                           \tiny{1:} &  Addition\_logistic(Matrix $A$,  Matrix $B$)  \\
                           \tiny{2:} &  SQL request for selecting $N$ number of processors   \\
                           \tiny{3:} &  Perform 2/1-D cyclic distribution of the Matrix $A$  $\vartriangleright$ Note: the distribution \\
                                         & is performed with the number of processors selected inside the SQL request. \\
                           \tiny{4:} &  Perform 2/1-D cyclic distribution of the Matrix $B$ \\
%                           \tiny{5:} &  Perform 2/1-D cyclic distribution of the Matrix $C$ \\ 
                           \end{tabular} & \\ \hline
		 \end{tabular}  
		 \caption{Logistic kernel associated to the matrix multiplication or  the addition.  \label{LKERNEL}}
	\end{center}
\end{table}

To illustrate, we describe the example of a simplified\footnote{without any C++ language annotations} logistic kernel of the matrix addition in table {\ref{LKERNEL}}. 
\begin{itemize}
\item Line 1:  the signature of the logistic kernel, showing the input/ouput by the keyword\texttt{const}. In this example there are onlz three matrices, but additional arguments like float number, integer \etc can be added.
\item Line 2: The SQL-like request selecting the desired number of processors for the operation and performing the required MPI calls. 
\item Line 3-5: Performing a block-cyclic distribution of the matrixes \texttt{A} and  \texttt{B}. A detailed description of the SQL requests and the distribution functions
is given in section \ref{SQLSECTION}. Note that the present kernel can be reused for any subsequent operation that uses the same data distribution.
\end{itemize}

\subsubsection{The computational kernel}  After the logistic kernel has been executed, we are ready for the computational kernel. The mode of execution is similar to the CUDA API with a few specificities. In the CUDA parallel computing model, the programmer writes a single thread program, using the CUDA-API,  after which the GPU runs multiple instances of this thread in parallel. We transpose this approach into Ambient. Nevertheless, we should not forget the data scope of a single CPU "thread" in Ambient roughly corresponds to the scope of a full OpenCL work group.

We have introduced a keyword (as the \texttt{const} keyword) named \texttt{pinned}. This keyword indicates to the framework than the kernel execution must be executed over all Ambient work groups. Inside a kernel, we employ several Ambient API, analogous to CUDA, to obtain key information about the work groups 
(positions, dimensions, \ldots).These are summarized in table \ref{ANALOGY}.  For illustration, the computational kernel (table \ref{CKERNEL}) of an addition of the two matrices associated with the previous example of the logistic kernel is presented in the table \ref{LKERNEL}.
\begin{itemize}
\item Line 1: the signature of the computational kernel must be the same than that of the associated logistic kernel. The only difference is the use of the keyword \texttt{pinned}, indicating an execution over all Ambient work groups.
\item Line 2: the cartesian coordinates of the work group are obtained from the Ambient instruction \texttt{get\_block\_id(A).x}  for $i$ and \texttt{get\_block\_id(A).y}  for $j$. \label{CURRENT}
\item Line 3 to 5 : we refer the work group $i,j$ to a temporary variable $A_{ref}$ using the Ambient instruction \texttt{current(A)(i,j)}. Note that if the processor requires data outside its memory,
 it will perform a remote access in the background (see section \ref{MPIABSTRACTION} for details).
\item Line 6 to 8 : the addition is performed by looping over the total size of the work group, in parallel for each workgroup (because of the use of the keyword \texttt{pinned}.
\end{itemize}

 \begin{table}[t] 
	\begin{center}
	 	\begin{tabular}{l l }
                           \hline 
                           \textbf{Computatiomal kernel }  Algorithm of the addition computational kernel  &    \\ \hline
                           \textbf{Require}                 A class Matrix  that fits with Ambient framework  &  \\
                           \begin{tabular}{c l} 
                               \tiny{1:} &  Addition\_computational(Matrix $A$,  \texttt{pinned const}  Matrix $B$) \\
                               \tiny{2:} &  Get the cartesian coordinate $i,j$ of the work group \\ 
                               \tiny{3:} &  Refer  the  work group$(i,j)$  of A into $A_{ref}$ \\
                               \tiny{4:} &  Refer  the  work group$(i,j)$  of B into $B_{ref}$ \\
%                               \tiny{5:} &  Refer  the  work group$(i,j)$  of C into $C_{ref}$  \\
                               \tiny{5:} &  \textbf{for} $k$ from $0$ to total size of the work group \textbf{do} \\
                               \tiny{6:} & \hspace{0.2 cm}   $A_{ref}[k] \, + \!\! = B_{ref}[k]$ \\
                               \tiny{7:} &  \textbf{end for} \\                               
                          \end{tabular} & \\ \hline
		 \end{tabular} 
		 \caption{Algorithm of the computational kernel associated to the matrix addition.   \label{CKERNEL}}
	\end{center}
\end{table}

As just shown for the matrix addition, we can recode other linear algebra solvers with the Ambient approach. For the presentation of our framework, we will focus on the MPI version of BLAS named pBLAS \cite{SCALAPACK-1992}, as well as a possible alternative to the pDGEMM solver, called ...., using a recursive algorithm.

%\begin{table}[h]
%	\begin{center}
%	 	\begin{tabular}{c  c  c}		 	                                                             & CUDA/OpenCL                                 & Ambient                               \\  \hline 
%			 work group-ID                                  &  \texttt{blockIdx}                                 & \texttt{get\_block\_id(A)}   \\  
%			 work group size                               &  \texttt{blockDim}                               & \texttt{get\_mem\_dim(A)}      \\  
%			 grid size                                            &  \texttt{gridDim}                                  & \texttt{get\_grid\_dim(A)}  \\ \hline
%				\end{tabular}
%		\caption{Analog¤y between Ambient and  the CUDA/OpenCL ideology. \texttt{A} is the associated matrix to the present logistic and computational kernel.  The access of the cartesian values are done by the classical operator C/C++ \guillemotleft.\guillemotright     \, \eg \texttt{get\_mem\_dim(A).x}  } 	
%		\label{ANALOGY2}
%	\end{center}
%\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END CUDA IDEOLOGY SECTION

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN SQL SECTION

\subsection{SQL resource management}  \label{SQLSECTION}
 In the previous section, we introduced the logistic kernels and their role in initializing resources for the execution of the computational kernels. 
 One of Ambients' main qualities is the ability to perform a dynamic management of the CPU resources of the cluster. To facilitate this, we integrated a light SQL framework \cite{SQLite} inside Ambient, which selects the desired number of processors inside the logistic kernel, in order to execute the computational kernel.

At the beginning of the framework execution, a main MPI group is created named \texttt{ambient}. To create groups inside this main group we used the abilities of MPI to build nested independent groups of processors  (chapter 6 of  \cite{MPI-2002}).  We encapsulated these features in the framework, and we control them through the SQL requests. This allows for a simple yet flexible management  of the MPI groups.  
These requests (done by the \texttt{scope\_select} Ambient-function) are applied to the logistic/computational addition kernels of our example by the following SQL syntax: \\

\begin{eqnarray}
       & &\texttt{scope\_select("\textit{N} from ambient as addition\_group }   \label{SQL1} \\
       & &\texttt{	where master is 0")}, \nonumber
\end{eqnarray}
This command selects \texttt{\textit{N}} processors from the original MPI group  \texttt{ambient}. The newly created group is 
named \texttt{addition\_group}, and its master processor locally has MPI-rank 0 (see Fig. \ref{MPIGROUP}).

\begin{figure}[b]
	\begin{center}
           	         \psfrag{Datas distributed over work group}{\tiny{\hspace{-0.5cm}Datas distributed over work group}}
	                  \psfrag{Local MPI rank}{\tiny{Local MPI rank}}                
	                  \psfrag{Ambient master proc}{\hspace{-0.8cm} \tiny{Ambient master proc}}        	                  
	                   \psfrag{Ambient MPI group, N process}{ \tiny Ambient MPI group, cardinal $N$ process}
	                   \psfrag{Local rank}{\tiny Local rank}
	         	         \psfrag{0}{\tiny{0}}
	                  \psfrag{1}{\hspace{-0.05cm}\tiny{1}}
	                  \psfrag{2}{\hspace{-0.15cm} \tiny{2}}
	                  \psfrag{MPI group creation}{ \hspace{-0.5cm} \tiny  Group creation}
	                  \psfrag{from the SQL request}{ \hspace{-0.6cm} \tiny  from SQL request}
	                  \psfrag{Mapping of the work group throw}{ \hspace{-0.6cm} \tiny  Mapping of the work group throw}
	                  \psfrag{the mpi GROUP created}{\tiny the new MPI group}
	                 \psfrag{(0,0)}{\hspace{-0.25cm} \tiny (0,0)}
	                 \psfrag{(0,1)}{\hspace{-0.25cm} \tiny (0,1)}
		        \psfrag{(0,2)}{\hspace{-0.25cm} \tiny (0,2)}
		        \psfrag{(1,0)}{\hspace{-0.25cm} \tiny (1,0)}
		        \psfrag{(1,1)}{\hspace{-0.25cm} \tiny (1,1)}
		        \psfrag{(1,2)}{\hspace{-0.25cm} \tiny (1,2)}
	                 \psfrag{(2,0)}{\hspace{-0.25cm} \tiny (2,0)}
		        \psfrag{(2,1)}{\hspace{-0.25cm} \tiny (2,1)}
		         \psfrag{(2,2)}{\hspace{-0.25cm} \tiny(2,2)}
                           \psfrag{n}{\hspace{-0.25cm} \tiny $n$}
	                  \includegraphics[scale=0.35]{FIGURES/MpiGroupData.eps} 
	\end{center}
          \caption[]{Creation of a MPI group by the SQL request, with \hbox{\cloud \hspace{0.02cm}} representing a MPI group with its own communicator;  $\bullet$ represents the master 
          process inside the group whereas the $\circ$ are the associated slave processors inside the corresponding group. Also shown is the assignment of the data blocks (each block corresponding to an Ambient work group) inside the newly created MPI group. In this example, the matrix is mapped, block by block, on the processors using a 1D column-cyclic distribution.}
          \label{MPIGROUP}
\end{figure}

After the SQL request, the data (workgroup) is mapped to the processes inside the newly created MPI group by a distribution function (\eg 1D or 2D block-cyclic distribution), the algorithm of which is described in table \ref{2DCYCLIC} (\cf Fig. \ref{MPIGROUP}). After this transfer, the processes of the MPI group are the primary owners in terms of data access and execution. In the Ambient approach, every operation is associated with
a logistic and a computational kernel, therefore the SQL request will be called several times. To create new groups, processes from the existing group are selected in ascending numerical rank, until all processes of the \texttt{ambient} group 
have been selected. If further groups are required, the selection is restarted from the beginning, thus a processor can belong to multiple groups simultaneously.

Every new group will have its own intra-communicator to allow intra-group communication, as well as an inter-communicator to allow for communication between groups, using the master process of the \texttt{ambient} group.

The construction of nested MPI-groups is related to the ability of MPI to create virtual and cartesian topologies  (chapter 7 of  \cite{MPI-2002}). However, Ambient allows for the simultaneous construction of the MPI topology and the mapping of the  data (work group).

 \begin{table}[t] 
	\begin{center}
	 	\begin{tabular}{l l }
                           \hline 
                           \textbf{Assign workgroup}  Algorithm of the 1/2-D assign function  &    \\ \hline
                           \textbf{Require}        An new MPI group obtained by an SQL request  &  \\
                           \begin{tabular}{c l} 
                               \tiny{1:} &  Initialize $i$ in function of the new MPI rank  of the new group \\
                               \tiny{2:} &  Initialize $j$ in function of the new MPI rank  of the new group  $\vartriangleright$ Note:   \\
                                             &  the initialization of $i,j$ will fix   the distribution (1D or 2D) \\
                               \tiny{3:} &  \textbf{for} from $i$  to the number of the work group on $x$ \textbf{do} \\
                               \tiny{4:} &  \hspace{0.2 cm}  \textbf{for} from $j$  to the number of the work group on $y$ \textbf{do} \\
                               \tiny{5:} & \hspace{0.4 cm}    Assign the work group $(i,j)$ to the corresponding processor\\
                               \tiny{6:} &   \hspace{0.2 cm}    \textbf{end for} \\
                               \tiny{7:} &  \textbf{end for} \\                               
                          \end{tabular} & \\ \hline
		 \end{tabular} 
		 \caption{Algorithm of the distribution of the work group into the new MPI group.   \label{2DCYCLIC}}
	\end{center}
\end{table}

During the execution, many group will be created in practice, and data will have to be moved between groups.
To obtain better performance and to reduce the number of data transfers between groups, there are additional possible settings for the SQL request. For example, consider the previous illustration for the SQL request in case of the matrix addition operation ($ A \, +\!\!\!= B$).
If matrix $A$ has already been created and mapped on specific processors, the logistic kernel can create a MPI group that includes its previous owner group (created in the previous operation), thereby avoiding superfluous data transfer. The concrete SQL request would be as follows:

\begin{eqnarray}  \label{SQL2}
         & & \texttt{scope\_select("\textit{N} from ambient as addition\_group }\\ 
         & & \texttt{	where master is 0 where breakdown contains" + get\_id(\textit{A})}, \nonumber
\end{eqnarray} 
The function \texttt{get\_id(\textit{A})} will get a tag  of \texttt{\textit{A}} to know is present binding.

%\begin{figure}[h]
%	\begin{center}
%			\includegraphics[scale=0.35]{FIGURES/ComGroup.eps} 
%	\end{center}
%          \caption{Data exchange inside the group.}
%          \label{COMGROUP}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END SQL SECTION

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGINNING EXECUTION AND OPERATIONS SCHEDULING SECTION
\subsection{Execution model and Operations scheduling}
In a typical HPC application, the programming model imposes a  parallel execution of every function of the code, with data granularity imposing the details 
of the parallelization. In Ambient, we developed an scheduler, an  \guillemotleft engine\guillemotright \, that privileges a queue execution model, where operations (\eg linear algebra operations \ldots)
are stacked up in a list first, and are executed afterwards according to the model of the logistics/computation kernel. This model is inspired by computer architecture design, i.e.
the queue data flow execution in a processor \cite{DENNIS-1974}. We are replicating this approach now on a larger scale where basic instructions ($+,-,*,/$) are replaced 
by the pairs of logistic/computational kernels.

Any HPC application can be decomposed into a set of functions performing numerical operations.  Starting from basic examples, we want to perform the three following operations :

\begin{eqnarray} 
\text{Operation 1:} & & A + \!\! =  B, \label{3OPERATIONS}  \\
\text{Operation 2:} & & A \times \!\!= \lambda,   \nonumber  \\ 
\text{Operation 3:} & & C \times \!\!= D. \nonumber
\end{eqnarray}
where $A$, $B$, $C$ and $D$ are some matrices and $\lambda$ is some scalar.

Conventionally, these three operations would be executed in parallel one after another. Transposing into Ambients' conception means associating each operation with a logistic and a computational kernel. Considering the first example of example (\ref{3OPERATIONS}),  a user would invoke the following \texttt{push} function:

\begin{eqnarray}
     \texttt{push(Addition\_logistic, Addition\_computional, \textit{A}, \textit{B})},
\end{eqnarray}

with the appropriate logistic and computational kernels \texttt{Addition\_logistic} and \texttt{Addition\_computational}  (c.f. table \ref{LKERNEL} and \ref{CKERNEL}), followed by the arguments \texttt{\textit{A}} and \texttt{\textit{B}}. In this way operations, such as those in (\ref{3OPERATIONS}), will stack up during execution of the application. At any time then, the operations on the stack can be executed using the basic Ambient instruction

\begin{eqnarray}
      \texttt{playout(),}
\end{eqnarray}

although ideally this command would be invoked at the end, when all operations have been pushed onto the stack.
The \texttt{playout()} command triggers the Ambient engine, scanning through the operation stack, identifying the independent operations and marking dependencies. Next, a dedicated MPI packet manager\footnote{a tool managing the communication between processors.} (see section \ref{MPIABSTRACTION})  is added for each operation, and all logistic kernels are executed. Where dependencies exist, the relevant data - which is mapped  to some MPI group according to the requirements of the first operation in which it is needed - will be marked as 'needed in other operations' ; those later dependent operations can always trace it to its then-current owner.
Finally, the computational kernels of independent operations are executed, followed by the dependent computational kernels in the appropriate order. In the case of data dependency, the data transfer is handled by the packet manager. When all computational kernels have been executed, the stack is cleared. A schematic of the stack execution is presented in Fig. \ref{EXECUTION} .

Overall, the Ambient approach combines two levels of parallelism: that of each computational kernel and that of several kernels executed in parallel by assigning them to different MPI groups.

\begin{figure}[h]
	\begin{center}
	      	        \psfrag{Operation 1}{\tiny{Operation 1}}
		        \psfrag{Operation 2}{\tiny{Operation 2}}
		        \psfrag{Operation 3}{\tiny{Operation 3}}
		        \psfrag{l/c kernel 1}{\tiny{l/c kernel 1}}
                           \psfrag{l/c kernel 2}{\tiny{l/c kernel 2}}
                           \psfrag{l/c kernel 3}{\tiny{l/c kernel 3}}
                           \psfrag{Abiemt}{\tiny{Ambient}}
                           \psfrag{schedule}{\tiny{Schedule}}
                           \psfrag{Dependency}{\tiny{Dependency}}
                           \psfrag{Time}{\tiny{Time}}
                           \psfrag{evolution}{\tiny{evolution}}
                           \psfrag{Mapping and execution}{\tiny{Mapping and execution}}                          
                           \psfrag{Ambient execution}{\tiny{Ambient execution}}                          
                           \psfrag{Usual execution}{\tiny{Usual execution}}                          
	                  \includegraphics[scale=0.35]{FIGURES/execution.eps} 
	\end{center}
          \caption{Model of execution of the ambient scheduler. On the classical approach (left) the operations will be executed in parallel step by step whatever the dependencies (marked by the symbol \guillemotleft [\guillemotright \, in the schematic). In ambient (right), operations will be associated to the pair logistics/computation (abbreviate l/c) kernel, as operations 1 and 3  (l/c kernels 1 and 3) are independents, the execution of computational kernel will be done, simultaneously.  Due to the dependency between the kernel 1 and 2, the computational kernel of operation 2 (l/c kernel 2) will be executed later. It should be performed in a new MPI group but  with an appropriate SQL request (\ref{SQL2}) in logistic kernel, we can execute the computational kernel where the data are. }   \label{EXECUTION}
\end{figure}

Next, we generalize example \ref{3OPERATIONS} to deal with vectors of matrices. We assume conditional statements determining which matrices should be added and/or multipled (Table \ref{GRAPH}). First, from line 1 to 11, as a dry run, the application execution will determine which operations should be performed, 
pushing the required kernels on the top of stack. When the \texttt{playout()} function is called line 12, the engine will execute the stack. By resolving dependency, it will create the directed acyclic graph of the application naturally. 
\begin{table}[t] 
	\begin{center}
	 	\begin{tabular}{l l }
                           \hline 
                           Determination of the directed  acyclic graph   &    \\ \hline
                           \textbf{Require}    A class Matrix  that fits with Ambient framework   &  \\  
                           and  vector of matrices  of size $n$, $A[n]$, $B[n]$, $C[n]$ and $D[n]$ & \\
                           \begin{tabular}{c l} 
                               \tiny{1:}   & \textbf{for} from $i$  to the size of the vector \textbf{do} \\
                               \tiny{2:}   & \hspace{0.2 cm} \textbf{for} from $j$  to the size of the vector \textbf{do} \\                               
                               \tiny{3:}   & \hspace{0.4 cm}  \textbf{if}  the conditions on $A[i]$ and  $B[j]$ are verified \textbf{do} \\                               
                               \tiny{4:}   & \hspace{0.8 cm}  $A[i]+ \!\! =B[j]$, $\vartriangleright$  push l/c kernel 1, \\
                               \tiny{5:}   & \hspace{0.4 cm} \textbf{end if} \\   
                               \tiny{6:}   & \hspace{0.4 cm} $A[i]\times \!\! =\lambda$,  $\vartriangleright$  push l/c kernel 2, \\                   
                               \tiny{7:}   & \hspace{0.4 cm} \textbf{if}  the conditions on $C[i]$ and  $D[j]$ are verified \textbf{do} \\                               
                               \tiny{8:}   & \hspace{0.8 cm} $C[i]\times \!\! =D[j]$,  $\vartriangleright$ push l/c kernel 3, \\
                               \tiny{9:}   & \hspace{0.4 cm}  \textbf{end if} \\       
                               \tiny{10:} & \hspace{0.2 cm}  \textbf{end do} \\   
                               \tiny{11:} & \textbf{end do} \\                                 
                               \tiny{12:} & \texttt{playout()} $\vartriangleright$  execute the stack \\
                               \tiny{13:} & $\vartriangleright$ Note:  $+\!\!=$ and $\times \!\! =$ operators have wrapper functions   \\
                                                & that push the good logistic and computational kernel. \\                               
                           \end{tabular}   &  \\ \hline
        	     \end{tabular}
	\end{center}
	\caption{Generalization of the example \ref{3OPERATIONS} with conditional statements. The association of a cumulative stack and  delate stack execution will determine the  directed  acyclic graph. \label{GRAPH}}
\end{table}

These examples demonstrate the strength of Ambient, allowing for two levels of parallelization without any hard coding of MPI instructions in the original code, with a great degree of flexibility and adaptability. Thus, a fully parallel linear algebra  application can be developed without taking into account the geometry of the problem, which usually done.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END EXECUTION AND OPERATIONS SCHEDULING SECTION

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN PACKET MANAGER 
\subsection{Packet manager and MPI abstraction} \label{MPIABSTRACTION}
In standard HPC applications, data may have to be prepared for exchange over the network, particularly if the data is of more than one type; in the latter case,
a MPI structure is set up, to which the data is then copied, and then this structure is send via network. These preliminary conversions can cut into
performance. In Ambient, the data layout inside each work group necessitates additional metadata to coordinate the execution of logistic and 
computational kernels, particularly the latter has to perform a remote access. Thus, before the block of data held by a workgroup, contiguous
metadata is included. We call this ensemble \guillemotleft block of data+metadata\guillemotright \, as a network packet, which is built
on a MPI structure. Thereby, work groups are able to send or receive without any additional movements or copying. 
To assure proper execution of the framework, several additional type of packets are  available, request packet, layout packet and data-block packet
they  all derived from a standard packet.

When a computational kernel is executed, it can access the any required, such as a block of matrix $A$, by use of the Ambient command  \texttt{current(\textit{A})(i,j)}
(ref. \ref{CURRENT}), where \textit{A} has to be a matrix  compatible with Ambient (\cite{ABSTRACTPARADATATYPE}), the indices $i$ and $j$ being the coordinates of the work group (Fig. \ref{MPIGROUP}). 
There are three possibilities for the resultant types of data communication when a process executes this function inside a computational kernel:

\begin{enumerate}
\item The work group is already in the memory of the process. The access is direct.
\item The work group is in the memory space of the present MPI-group (previously created by the associated logistic kernel). 
The process requiring this data sends a query (a specific type of packet) to the master process of its MPI group.
The master process, which knows which slave process of its' group holds the required work group, sends a new request to
the that process, which in turn sends the requested workgroup to the requesting process. 
\item The work group is held inside another MPI-group. As previously, the process which executes the computational kernel sends a query to the master process of its MPI group,
which in turn sends a request to the master process of the MPI group within which the data is held. That master process then sends a message to the actual owner process,
which finally sends the data to the initially requesting process.
\end{enumerate}
Case (2) and (3) are illustrated in Fig. \ref{MPIEXCHANGE}.
\begin{figure}[h]
	\begin{center}
	    \psfrag{case 2}{\tiny case 2}
	    \psfrag{case 3}{\tiny case 3}
             \includegraphics[scale=0.35]{FIGURES/ComGroup.eps} 
	\end{center}
          \caption{On the left: communications inside the same group. On the right: communications between two MPI-groups. The dashed circle indicates the group holding the requested data.
          In both cases, the master processes have a key role in distributing information for the exchange. Each arrow marks a request, the sequence of requests is enumerated.}  \label{MPIEXCHANGE}
\end{figure}

Clearly, smooth execution of any communication pattern arising from the great number of possible data dependencies
is key to a good performance of the Ambient approach. The resolution of communication patters also falls under the purview
of the packet manager, as it manages both intra- and inter-group communication. Each MPI-group has a unique packet manager,
identified by a singleton pattern.

To reach the best performance in HPC, bottlenecks arising from synchronization operations should be avoided  \cite{EXASCALE-2010}.
Thus, MPI asynchronous communication is privileged in Ambient (\texttt{mpi\_isend} and \texttt{mpi\_irecv}). Still, an asynchronously
sent message  requires a \texttt{mpi\_test} to check whether the packet can be received.  As we avoid synchronization instructions,
the receiving process repeatedly checks for the availability of the expected data in an infinite loop, until actually receiving the
data.

 For the packet manager to be adaptable to any communication pattern, it consists of two parts:
the delegate pattern and a finite state machine (FSM), both are placed inside this loop.

\begin{table}[b] 
	\begin{center}
	 	\begin{tabular}{l l }
                           \hline 
                           Algorithm of the packet manager   &    \\ \hline
                           \textbf{Require}  a MPI process require a block of data &  \\                           
                            \begin{tabular}{c l}                                
                               \tiny{1:}   & \textbf{for}  ;;  \textbf{do} $\vartriangleright$  infinite loop \\
                               \tiny{2:}   & \hspace{0.2 cm}  \texttt{Delegate()}  $\vartriangleright$ check if  packet has been, \\
                                                & and received and re-send if needed \\
                               \tiny{3:}   & \hspace{0.2 cm}  FSM $\vartriangleright$ break the loop, \\
                                               & if all packets arrived at destination. \\
                               \tiny{4:}   & \textbf{end do} \\
                           \end{tabular}   &  \\ \hline
        	     \end{tabular}
	\end{center}
	\caption{Description of the packet manager}
\end{table}

%\begin{figure}[h]
%	\begin{center}
%	    \psfrag{1}{\tiny \hspace{-0.1 cm}  1}
%	    \psfrag{2}{\tiny \hspace{-0.1 cm}  2}
%	    \psfrag{Loose}{\tiny \hspace{-0.2 cm}   Loose}
%	    \psfrag{Closure}{\tiny \hspace{-0.2 cm}  Closure}
%	    \psfrag{Open}{\tiny \hspace{-0.4 cm}  Opening}
%          \includegraphics[scale=0.45]{FIGURES/FSM.eps} 
%	\end{center}
%          \caption{Schematic of the finite sate machine for the closing procedure. The states 1 and 2 correspond to open or close.
%           The transitions are loose, closure or opening, they  depend of the situation of all packets (if they all reach theirs targets or not),
%                          the loose transition means the packet has not been received, thus we iterate one more time on the infinite loop.}  \label{FSM}
%\end{figure}

The delegate pattern (from C\#) encapsulates  methods, behaving similar to a function pointer in C++,  but allowing for safer execution.
The delegate object inside the packet manager class contains a series of functions pile up in a stack, such as forwarding requests, 
layout and block of data. As this delegate is inside the infinite loop, the delegate execution 
will propagate the packets process by process after every iterations (as shown on Fig. \ref{EXECUTION}). When the delegate execution is over, a finite state machine
checks the states of the packets (arrived or not).  If all packets (data packet, layout packet, etc.) have been received, the FSM breaks the infinite loop.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END PACKET MANAGER

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN ABSTRACT PARALLEL DATATYPES 
\subsection{Abstract parallel datatypes} \label{ABSTRACTPARADATATYPE}

To assure the largest scope to Ambient, we designed the framework on the data structures and not a specific problem. Thus, we enlarge the utilization of the framework
for a various set of applied computing fields. The connection  between Ambient and the \guillemotleft outside world \guillemotright  \, is done by an intermediate class that assure the
connectivity between a class and ambient.

For linear algebra, we designed a matrix class, we contained usual data members for describing the object (number of raws, number of columns, \dots) and  methods for linear
algebra operations (associated to logistic and computational kernel).This matrix class derived from an other who assured the connection to Ambient.
When an instance of the matrix class is created, the inheritance mechanics will call constructors, methods in the mother class that  will 
structure  the memory (tile  layout).

Therefore, we can extend Ambient for new problems. Per example, in fluid mechanics the  two-dimensional finite difference method is very popular for solving the Navier-Stokes  differential equations.
These equations describes the parameters of the \guillemotleft fluid particle \guillemotright. In this case Ambient will discretize the space and assure, as for a matrix,  the distribution
 of the data inside the memory. Thus, a fluid particle class will contain enough information to assure a good discretization in the cartezian space (as the number of raws and columns inside the matrix class).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END ABSTRACT PARALLEL DATATYPES

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN VALIDATION
\subsection{Validation}
As a validation test, we code a new implementation of PDGEMM using the Ambient framework, and compare to the ScaLAPACK version of PDGEMM (MKL-Version).
The algorithm of our implementation is based on the classical method [CITATION NEEDED] using a column order distribution for the matrices (\cf table \ref{AMBIENTDGEMM} 
for the modification due to Ambient). The main advantage in our implementation is the required amount of coding, which is limited to seven lines.

----------------------------------------------------------------------------------------------------- \\
TO DO: EXPLAIN THE ADVANTAGE AGAINST   SCALAPACK \\
IN TERM OF NETWORK, DATA EXCHANGE \\
----------------------------------------------------------------------------------------------------- \\

We compare the Ambient results with the ScaLAPACK implementation. As related in section \ref{CUDAIDEOLOGYSECTION}, Ambient is compatible with ScaLAPACK,
despite the difference in memory layout. In Ambient, each work group is principally independent from any other, even if they belong to the same process, while in ScaLAPACK
blocks are grouped by process s.t. a memory-contiguous object is obtained. Therefore, calling a ScaLAPACK solver from Ambient must entail rearrangement of appropriate
work groups into a contiguous blocks, and thus a certain overhead. Due to Ambients flexibility in handling and grouping data, from a programming perspective this task
is simple to accomplish however.

 \begin{table}[t] 
	\begin{center}
	 	\begin{tabular}{l l }
                           \hline 
                           \textbf{ Algorithm:}  pDGEMM using Ambient  &    \\ \hline
                           \textbf{Require}                 A pinned computational kernel  &  \\
                           \begin{tabular}{c l} 
                                 \tiny{1:} &  Get the cartesian coordinate $i,j$ of the work group \\ 
                                 \tiny{2:} &  Refer  the  work group$(i,j)$  of B into $B_{ref}$ \\
                                 \tiny{3:} &  \textbf{for} from $k(0)$  to the max number of rows \textbf{do} \\
                                 \tiny{4:} &  \hspace{0.2 cm} Refer  the  work group$(k,j)$  of A into $A_{ref}$ \\    
                                 \tiny{5:} &  \hspace{0.2 cm}  reduction of the blocks $(k,i)$ into $C_{ref}$ , $\vartriangleright$  The reduction, \\
                                                & \hspace{0.2 cm}  is done by a proxy pattern \\
                                 \tiny{6:} &  \hspace{0.2 cm}  classical DGEMM($A_{ref}$, $B_{ref}$, $C_{ref}$)\\
                                 \tiny{7:} &  \textbf{end do} \\    
                             \end{tabular} & \\ \hline
		 \end{tabular} 
		 \caption{Algorithm of the pDGEMM using Ambient framework. We remember a proxy pattern is an interface the something else (object in memory). There, it allows the reduction between the independent workgroups.   \label{AMBIENTDGEMM}}
	\end{center}
\end{table}

The validation test consists of multiplying matrices and saving the results into an output matrix, $C = A \times B$,
where we further compare between the output of the Ambient computational kernel and the ScaLAPACK PDGEMM solver.
Using the C++ language, the operations $\times$ and $=$ have been overloaded appropriately, now containing the "pushing" of the required logistic and computational kernels.
Afterwards, results are compared element by element using the test  \cite{Dongarra:1990:SLB:77626.79170}  $\left \lvert C^S_{ij} - C^A_{ij} \right \rvert/\left \lvert  \epsilon C^S_{ij} \right \rvert  < n$,
where  the precision factor $\epsilon$  is the \texttt{double} precision $10^{-16}$, indices  $S$ and $A$ represent the ScaLAPACK and the Ambient version, and $n=16$, the exponent of numerical precision.

For the test, the size of work groups has been fixed to $128 \times 128$, and the square matrices $A$, $B$ range from 256 up to 8196 in size (step size of 128). The numbers of MPI-Process varied from 2 to 8.
Matrices were filled upusing a pseudo random number generator  \texttt{drand48()}\footnote{This function returns 
pseudo-random numbers non-negative, double-precision,  uniformly distributed over the interval $[0, 1]$ using a linear congruential algorithm and 
48-bit integer arithmetic. }  from \texttt{stdlib}.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END VALIDATION

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END DESCRIPTION OF THE FRAMEWORK

\section{Results and discussions}
\subsection{First benchmark case}
[STARTED REWRITING THIS; BUT NOT FINISHED]
The first benchmark consists of evaluating the performance of our parallel matrix multiplication relative to the iScaLAPACK-MKL mplementation. Both solvers are included as computational kernels
in Ambient.  The initial condition are the same than the validation test. As the validation test, the matrices size vary from 256 up to 8196 by step of (128 elements). A 1-D cyclic  column distribution is
used for both algorithm. As both kernels are included inside the framework additional workload should take into account and for the ScaLAPACK pDGEMM, we reorganize data to fit the ScaLAPACK
representation. The benchmarks  were performed on AMD Opteron  Magny-Cours) CPUs at 2.1 GHz on the Swiss National Supercomputing Centre facilities. The performance is reported in [GFlops], it is simply
calculated by :

\begin{eqnarray}
P  = \frac{2n^i}{t},
\end{eqnarray}
where $n$ is the size of the square-matrix,  $t$ the full time of calculation and $i$ is equal to 3 for the classical matrix multiplication algorithm.

\begin{figure}
\include{PLOT/SinglePDGEMM/SinglePGEMM}
\caption{Performances of PDGEMM using ScaLAPACK }
\end{figure}

\subsection{Second benchmark case}
This benchmarks consists to evaluate the performance of the scheduler, especially the repartition of several operation all over the cluster.

\begin{figure}
\include{PLOT/VectorPDGEMM/VectorPGEMM}
\caption{Performances of a vector of PDGEMM using ScaLAPACK}
\end{figure}

\subsection{Third benchmark case}

\section{Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN THANK 
\thanks{The authors are grateful to W. Sawyer and G. Fourestey from the CSCS staff and V. Rezzonico from EPFL.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END THANK

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN BIBLIO
\bibliographystyle{abbrv}
\bibliography{biblio}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END BIBLIOr

\end{document}