### 2013-05-16 doc/ambient/tensor_ws/tikz/energy_contraction.tex
tikz diagram for ambient poster

\documentclass{standalone}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{color}
\usepackage{amsthm}

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,shapes,positioning,decorations.pathmorphing}

\begin{document}

\tikzset{spin/.style={circle=2pt,draw=black!100,fill=blue!20,inner sep=3pt}}
\tikzset{hamil/.style={rectangle=2pt,draw=black!100,fill=red!40,inner sep=3pt}}
\tikzset{coarse grain/.style={isosceles triangle,isosceles triangle stretches,draw, fill=green!15, anchor=center, inner sep=3pt, minimum width=.8cm, shape border rotate=270, thick}}
\tikzset{arrow down/.style={single arrow, draw, fill, shape border rotate=270, anchor=center, minimum height=.4cm, single arrow head extend=.2cm, inner sep=8pt}}

\begin{tikzpicture}

\foreach \i in {1,...,5} {
	\node[spin] (M1 \i) at (\i,0) {};
%	\draw[thick] (M1 \i.south) -- +(0,-.4);
	\node[hamil] (H \i) at (\i,-.5) {};
	\draw[thick] (M1 \i) -- (H \i);
	\node[spin] (M2 \i) at (\i,-1) {};
	\draw[thick] (H \i) -- (M2 \i);
}
\foreach \i [evaluate = \i as \j using int(\i+1)] in {1,...,4} {
	\draw[thick] (M1 \i) -- (M1 \j);
	\draw[thick] (H \i) -- (H \j);
	\draw[thick] (M2 \i) -- (M2 \j);
}

%\draw[dashed,thick] (M1 5.east) -- +(.4,0);
%\draw[dashed,thick] (M1 1.west) -- +(-.4,0);

\end{tikzpicture}

\end{document}