### 2014-04-01 doc/ambient/tensor_ws/tensor_ws.tex
Initial version, svn r4227

\documentclass[a4paper]{article}

%% do not edit the following macro
\usepackage{graphics}
\newcommand{\abs}[4]{
\begin{center}
  {\large\bf #1}

   \vspace*{.6em}
  {#2}

   \vspace*{0.2em}
  {\it #3}

\end{center}
{#4}
}

\begin{document}
\abs{%
% 1. Title of Your Talk
Exposure of Dataflow parallelism in Density Matrix Renormalization Group}{%
%%%% 2. List of Authors
Alexandr Kosenkov$^1$, Michele Dolfi$^1$, Timothee Ewart$^2$, Adrian Kantian$^{2}$, Bela Bauer$^{3}$, Matthias Troyer$^1$}{%
%%%% 3. List of Affiliations
$^1$Theoretische Physik, ETH Zurich, 8093 Zurich, Switzerland\\
$^2$DPMC-MaNEP, University of Geneva, 1211 Geneva, Switzerland\\
$^3$Microsoft Research, Station Q, University of California, Santa Barbara, CA 93106, USA}{%
%%%% 4. Abstract

\begin{itemize}
\item Density Matrix Renormalization Group application deals with tensor operations using complex matrices objects. The application is automatically parallelized using Ambient: dataflow distributed memory framework for linear algebra operations.
\end{itemize}

}

\end{document}